---
title: "NativeModel Results."
date: '`r format(Sys.time(), "%d %B, %Y (%H:%M)")`'
output:
  html_document: 
    toc: yes
    pandoc_args: [
      "+RTS", "-K64m", 
      "-RTS"
    ]
fontsize: 10pt
---

# ResultsSummary - **NativeModel**

Analyses the results of the keanu tests when experimenting with parameter estimation (probabilistic model needs to guess a single threshold value ). Relevant files that generate the results: `NativeModel`

# Initialise and load data

```{r initialise, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(gtools)
library(ggplot2)
library(hexbin) # (this is called automatically if required, but putting it here ensures it has been installed)
library(gridExtra) # For arranging two grids side by side
#library(reshape2) # for melting (taking wide format and making it narrow)
library(xtable) # for making nice tables


homedir<-"C:/ABM_Uncertainty/dust/Improbable/NativeModel-keanu/results/"
setwd(homedir)

# Find files
samples.files <- grep("Results.*csv", list.files(path = "."), value=TRUE)
params.files <- grep("Params.*csv", list.files(path = "."), value=TRUE)

if (length(samples.files) != length(params.files)) {
  stop("For some reason there are a different number of results files and random number files.
       This is weird so I'm not continuing.")
}

if (length(samples.files) != 1) {
  message("Found multiple samples files. Using the most recent one.")
  samples.files <- tail(sort(samples.files), n=1)
  params.files  <- tail(sort(params.files),  n=1)
}

# Read the file in
samples <- read_csv(samples.files[1], col_names=FALSE)
params  <- read_csv(params.files[1])

samples[,ncol(samples)] <- NULL # Last column is NA because I lazily write too many commas
params[,ncol(params)] <- NULL # Last column is NA because I lazily write too many commas

# Truth data are the first line
truth.results <- samples[1,]
samples <- samples[-1,]

# Truth data are the first entry in the "Threshold" column
truth.params <- params[1,"Threshold"]
params <- params[-1,"Threshold"]

```

# Plot results (look at the people per iteration)

## Spaghetti plot

```{r spag.plot, fig.width=11, fig.height=9}
# Define a fucntion to plot the samples
spag_plot <- function(samples, truth) {

  x_axis <- 1:ncol(samples)
  sample_col <- alpha("#ADD8E6", 0.9)
  truth_col <- "red"

  #create plot
  plot(1, xlim=c(0, max(x_axis)), ylim=c(min(rbind(truth,samples)), max(rbind(truth,samples))), type='l',
       xlab="Number of iterations", ylab = "Number of agents" )
  map(1:nrow(samples), function(x) lines(x_axis, samples[x,], type='l', col=sample_col))
  lines(x_axis, truth, type='l', lwd=2.5, col = truth_col)
  abline(v = c(seq(0, 2000, by=200)), lty=2, col='green4')

  legend("bottomright", c("Sample", "Truth"), lty=c(1,1), lwd=c(2.5,2.5),col=c(sample_col, truth_col))
}


#map2(samples, obIntervals, function(x, obInterval) spag_plot(x, truth, obInterval))
spag_plot(samples, truth.results)

```


## Density plot

Create the plots

```{r density.plot-create}

# Returns a single ggplot2 density plot
density_plot <- function(s, truth) {

  # Y values for this sample (num people per iteration)
  y <- t(s[1,])[,1]
  # X values
  x <- 1:length(y)
  # A long two-column matrix will have number of people at every iteration across all samples
  scatter <- cbind(x, y)
  # Now add the rest of the samples
  for (i in 2:nrow(s)) {
    scatter <- rbind(scatter, cbind(1:length(y), t(s[i,])[,1]) )
  }
  scatter <- as.data.frame(scatter)

  the.plot <- ggplot(data=scatter, mapping=aes(x,y)) +
    geom_hex(bins=35, show.legend = FALSE) +
    geom_line(data = data.frame("x"=1:length(truth), "y"=t(truth)), mapping=aes(x,y), colour="red") + # Truth
    #geom_smooth(method="loess", se=TRUE, level=0.95, color="orange") +
    #geom_point(size=0.1, color="white") + # individual points
    ylab("Number of agents in the system")+
    xlab("Iteration")
    #ggtitle(paste("Observation Interval =", obs, sep =' '))

  return(the.plot)
}

#plots <- map2(samples, obIntervals, function(x, obInterval) density_plot(x, truth, obInterval))
density_plot(samples, truth.results)

```

# Plot parameters

Look at how the threshold parameter changed.

Start with a histogram.

```{r randPosterior, fig.width=7, fig.height=5}
param_hist <- function(params, truth.params) {
  
  x_range <- unlist(c(truth.params[1[[1]]], params[,1][[1]]))
  x_min <- min(x_range)
  x_max <- max(x_range)
  
  x_lim <- c(x_min-.2, x_max+.2)
  
  plot(density(params[,1][[1]]), col='blue', main='Threshold parameter posterior', xlim=x_lim)
  hist(params[,1][[1]], xlim=c(-1.1,1.1), prob=TRUE, breaks='Scott', add=TRUE, border='gray')
  abline(v=truth.params[1[[1]]], col='red')
  
  legend('topright', legend=c('True threshold value', 'Posterior'), lty=c(1,1), col=c('red', 'blue'))
}

param_hist(params, truth.params)
```

## 'True' parameter value(s)

```{r randTruthTable, results="asis"}
print(xtable(
  truth.params
  ), type = "html")
```

## Parameter samples

```{r randTable, results="asis"}
print(xtable(
  params
  ), type = "html")
```



