\section{ODD Protocol}
\label{sec:odd}

ODD (Overview, Design Concepts, Details) is a protocol for outlining and
documenting agent-based models
\cite{grimm2006standard,grimm2010odd,grimm2020odd}.
The aim of such a protocol is to allow researchers to share and reproduce their
models via a standardised format.
The protocol consists of three components, each of which cover a number of
different topics:
\begin{enumerate}
    \item Overview providing an overview of the model, touching upon each of the
        following topics outlined in Section \ref{sub:odd:overview}.
    \item Design Concepts: detailing which of the following design concepts are
        relevant to the model (detailed in Section
        \ref{sub:odd:design_concepts}).
    \item Details: elaborating on the internal mechanics of the model, which
        will cover each of the topics outlined in Section \ref{sub:odd:details}.
\end{enumerate}

\subsection{Overview}
\label{sub:odd:overview}

The Overview section of ODD aims to provide a cursory overview of the model that
is being documented.
This includes subsections on each of the following topics:
\begin{enumerate}
    \item Purpose and patterns: What can a user  expect the model to do/achieve,
        and what patterns might the model be fit to simulation
        \cite{grimm2012pattern}?
    \item Entities, state variables and scales: What types of agents and
        environments are we looking to model, what variables are we using to
        represent them, and what spatial and temporal scales are we working at?
    \item Process overview and scheduling: What processes govern the
        interactions in the model, and in what order to they occur?
\end{enumerate}
This information should enable others to develop and outline of the what the
model looks like at a functional level.

\subsection{Design Concepts}
\label{sub:odd:design_concepts}

When designing an agent-based model, there are a number of design concepts that
are often incorporated.
The following is a non-exhaustive list of the design concepts that may be
involved in an agent-based model:
\begin{enumerate}
    \item Basic principles
    \item Emergence
    \item Adaptation
    \item Objectives
    \item Learning
    \item Prediction
    \item Sensing
    \item Interaction
    \item Stochasticity
    \item Collectives
    \item Observation
\end{enumerate}
These design concepts are detailed in greater depth in \cite{grimm2010odd}.

\subsection{Details}
\label{sub:odd:details}

The Details section of ODD aims to build upon the above sections, providing the
specifics required to build a fully functioning model.
This takes the form of the following topics:
\begin{enumerate}
    \item Initialisation: What information is required to start the model with
        respect to defining the initial model state and calibrating parameter
        values?
    \item Input data: What information does the model require as it runs?
    \item Submodels: How do the individual processes governing model behaviour
        work?
\end{enumerate}

