\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Cover letter}
%\author{Le Minh Kieu, Dong Ngoduy, Nasser Sabar and Edward Chung}
\date{$17^{th}$ June 2019}

\begin{document}

\maketitle

Dear Editor,

Enclosed is the revised manuscript and response to reviewers of the manuscript entitled "Dealing with uncertainty in
agent-based models for
short-term predictions". Please accept it as a candidate for publication in the Royal Society Open Science. 

Agent-based models (ABM) are gaining traction as one of the most powerful modelling tools within the social sciences.
They are particularly suited to simulating complex systems. Despite many methodological advances within ABM, one of the major drawbacks is their inability to 
incorporate real-time data to make accurate short-term predictions. This paper presents an approach that allows ABMs to be dynamically optimised. Through a combination of parameter calibration and data assimilation (DA), the accuracy of model-based predictions using ABM in real time is increased.  We use the exemplar of a bus route system to explore these methods.  This system is spatially and temporally dynamic with the bus route conditions changing during peak periods (e.g. traffic) and complexity introduced by the varying numbers of passengers using the service at each stop.

Within this paper we construct an ABM that simulates the main interactions and key processes within this system.  To enable accurate real time predictions, we dynamically optimise the ABM through a combination of parameter calibration and DA techniques.  Finally, we quantify the impact of calibration and DA in dealing
the with stochastic and dynamic nature of the system under study.

Finally, this paper is our original work and has not been submitted anywhere else. We are looking forward to the constructive feedback from the anonymous reviewers on this work. 

Cordially,

Le-Minh Kieu, Nicolas Malleson and Alison Heppenstall. 

\end{document}
